---
title: "Data cleaning and figures"
output: html_document
date: "2025-05-15"
---

Project set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Set working directory
setwd("/Users/zoe/Desktop/CCBP/CCBP_database")

# Load required libraries and install packages if necessary
install.packages("ggalluvial")
install.packages("ggrepel")
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(ggalluvial)
library(readxl)
library(countrycode)
library(ggrepel)
library(dplyr)
library(countrycode)


```

Set up dataframe
```{r}
#Read in csv
# Data <- read.csv("C:/Users/uqbwil14/Dropbox/Carbon crossing borders project/Data from NTU (Singapore)/Carbon_Market_Data_20250408.csv")

# reading in new updated and combined dataset
load("Combined_Carbon_Market_Data_Expanded.RDATA")
Data <- combined_data_expanded

head(Data)

# Define keywords that indicate forest-based carbon projects
forest_keywords <- c("REDD", "redd", "Forest", "forest", "FOREST", "Reforestation", "REFORESTATION", "reforestation", "Afforestation", "afforestation", "AFFORESTATION", "Deforestation", "deforestation", "DEFORESTATION", "Ecosystem Conversion", "ecosystem conversion", "ECOSYSTEM CONVERSION", "Peat Swamp", "peat swamp", "PEAT SWAMP", "Forestry", "forestry", "FORESTRY", "tree", "Tree", "trees", "TREES", "TREE", "forestation", "Forestation", "FORESTATION", "planting", "Planting", "PLANTATION", "mangrove", "Mangrove", "MANGROVE", "swamp-forest", "swamp forest", "SWAMP FOREST", "Swamp Forest", "Swamp forest", "wood", "Wood", "WOOD", "regeneration", "Regeneration", "REGENERATION", "regrowth", "Regrowth", "REGROWTH", "vegetation", "VEGETATION", "Vegetation", "seedling", "Seedling", "SEEDLING", "sapling", "Sapling", "SAPLING", "bosque", "Bosque", "BOSQUE", "forestal", "Forestal", "FORESTAL", "plantaciones forestales", "PLANTACIONES FORESTALES", "arbol", "Arbol", "ARBOL", "Agro-forestry", "agro-forestry", "AGRO-FORESTRY", "agroforestry", "Agroforestry", "AGROFORESTRY", "Plantation", "plantation", "PLANTATION", "woodland", "Woodland", "WOODLAND")


#### NOTE TO BROOKE -- I changed the column name to match the new dataset column name for methodologies #####

# Identify forest-based projects -- only for methodologies - ZEL expanded to project_name and project_description in code below

# forest_projects_df <- Data[grepl(paste(forest_keywords, collapse = "|"), 
                           #    Data$project_methodologies, ignore.case = TRUE), ]


### ZEL - Expand filter to include more than just methodologies column for keyword filter - added project_name and project_description

forest_projects_df <- Data %>%
  filter(
    str_detect(project_methodologies, regex(paste(forest_keywords, collapse = "|"), ignore_case = TRUE)) |
    str_detect(project_name, regex(paste(forest_keywords, collapse = "|"), ignore_case = TRUE)) |
    str_detect(project_description, regex(paste(forest_keywords, collapse = "|"), ignore_case = TRUE))
  )

forest_projects_by_registry <- forest_projects_df %>%
  count(registry, sort = TRUE)

forest_projects_by_registry

## seems to work! All of the registries I expected to see are here.

# Print the result
print(forest_projects_df)

#### HI BROOKE - also need to change some column headings here. I think Headquarter.Country you want to be the proponent country? As in, the country funding the project but not necessarily where the project happens. I'm going on this assumption but let's change if I have gotten this wrong.


# Ensure Sum.of.Credit.Volume is numeric (cleaning empty or non-numeric entries)
## Data_reduced <- forest_projects_df %>%
#   select(
#    Headquarter.Country,
#    Buyer.Continent,
#    Credit.Volume = `Sum.of.Credit.Volume`,
  #  Project.Country,
   # Project.Continent
#  )


Data_reduced <- forest_projects_df %>%
  select(
    proponent_country_manual,
    buyer_continent,
    sum_of_credit_volume,
    host_country_or_area,
    project_continent,
    registry,
    project_methodologies,
    source
  )


# Filter for complete cases (no NA or empty strings)
# Data_complete <- Data_reduced %>%
#  filter(
#    complete.cases(.) &                     # Remove NA values
#    Credit.Volume != "" &                   # Remove empty strings
#    Headquarter.Country != "" &
#    Buyer.Continent != "" &
#    Project.Country != "" &
#    Project.Continent != ""
#  )

# Filter for complete cases (no NA or empty strings)
Data_complete <- Data_reduced %>%
  filter(
    complete.cases(.) &                     # Remove NA values
    sum_of_credit_volume != "" &                   # Remove empty strings
    proponent_country_manual != "" &
    buyer_continent != "" &
    host_country_or_area != "" &
    project_continent != ""
  )

head(Data_complete)
print(Data_complete)


save(Data_complete, file = "ccbp_market_data_forest_based.RDATA")


##### summarising methodologies into a table #####

method_summary <- Data_complete %>%
  group_by(registry) %>%
  summarise(
    n_methods = n_distinct(project_methodologies),
    methods   = paste(sort(unique(project_methodologies)), collapse = "; "),
    .groups = "drop"
  ) %>%
  arrange(registry)

method_summary

# making a table

install.packages("flextable")
install.packages("officer")

library(flextable)
library(officer)

# Build the flextable
ft <- flextable(method_summary)

# Autofit registry + count, but wrap methods column
ft <- autofit(ft) %>%
  width(j = 3, width = 4) %>%   # make the methods column 4 inches wide
  set_table_properties(layout = "autofit") %>%
  theme_vanilla() %>%
  add_header_lines("Summary of unique forest project methodologies per registry") %>%
  align(align = "left", part = "all")   # left-align wrapped text

# Save to Word
doc <- read_docx() %>%
  body_add_flextable(ft) %>%
  body_end_section_landscape()   # optional: makes it landscape if table is wide

print(doc, target = "forest_methods_by_registry_try4.docx")

###--------------------------------------------
###--------------------------------------------

##### Finding duplicates in methodologies #####

# Load required packages
library(dplyr)
library(stringr)
library(tidyr)
library(flextable)
library(officer)

method_summary_clean <- method_summary %>%
  rowwise() %>%
  mutate(
    # split into a list of methods per registry
    methods_vec = list(str_split(methods, ";\\s*")[[1]] %>% str_trim()),
    # remove duplicates + sort
    methods_unique = list(sort(unique(methods_vec))),
    # recombine into string
    methods = paste(methods_unique, collapse = "; "),
    # count unique
    n_methods = length(methods_unique)
  ) %>%
  ungroup() %>%
  select(registry, n_methods, methods) %>%
  arrange(registry)

# --- make nicely wrapped table for Word ---
ft <- flextable(method_summary_clean) %>%
  autofit() %>%
  width(j = 3, width = 4) %>%   # wrap 3rd column (methods)
  set_table_properties(layout = "autofit") %>%
  theme_vanilla() %>%
  add_header_lines("Summary of unique forest project methodologies per registry (deduplicated)") %>%
  align(align = "left", part = "all")

ft
# --- export to Word file ---
doc <- read_docx() %>%
  body_add_flextable(ft) %>%
  body_end_section_landscape()   # optional: remove if you prefer portrait

print(doc, target = "forest_methods_by_registry_deduplicated.docx")
```

```{r}
##### Cleaning ccbp_market_data_forest_based.RDATA to fix continents, ISO3 codes, and clearing up mistaken spellings of countries

##### add-on to reassign continents ######
#### Step 0 Define country groups by ISO3 ####

library(dplyr)
library(countrycode)

##### Step 0 — Define country groups by ISO3 (modern) ####

north_america_iso3 <- c("USA", "CAN")

south_america_iso3 <- c(
  "ARG", "BOL", "BRA", "CHL", "COL", "ECU",
  "GUY", "PRY", "PER", "SUR", "URY", "VEN"
)

caribbean_iso3 <- c(
  "ATG", # Antigua and Barbuda
  "BHS", # Bahamas
  "BRB", # Barbados
  "CUB", # Cuba
  "DMA", # Dominica
  "DOM", # Dominican Republic
  "GRD", # Grenada
  "JAM", # Jamaica
  "KNA", # Saint Kitts and Nevis
  "LCA", # Saint Lucia
  "VCT", # Saint Vincent and the Grenadines
  "TTO",  # Trinidad and Tobago
  "HTI"
)

central_america_iso3 <- c(
  "MEX", "GTM", "BLZ", "HND", "SLV", "NIC", "CRI", "PAN", "HTI",
  caribbean_iso3
)

# Sanity check of country sort

host_carbon_wgi %>%
  mutate(
    region = case_when(
      iso3c %in% north_america_iso3   ~ "North America",
      iso3c %in% central_america_iso3 ~ "Central America & Caribbean",
      iso3c %in% south_america_iso3   ~ "South America",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(is.na(region)) %>%
  distinct(iso3c, countryname)


##### Step 1 — Country-name cleaning helper ####

clean_country_name <- function(x) {
  case_when(
    x %in% c("US", "United States") ~ "United States of America",
    x %in% c("MX")                  ~ "Mexico",
    x %in% c("Republic of Moldova", "Moldova") ~ "Moldova",
    TRUE ~ x
  )
}

##### Step 2 — Clean names + assign ISO3 (modern + legacy) ####

Data_complete_iso <- Data_complete %>%
  mutate(
    host_country_clean      = clean_country_name(host_country_or_area),
    proponent_country_clean = clean_country_name(proponent_country_manual),

    # modern ISO3 (for geography)
    ISO3_host_modern = countrycode(
      host_country_clean, "country.name", "iso3c"
    ),
    ISO3_proponent_modern = countrycode(
      proponent_country_clean, "country.name", "iso3c"
    ),

    # legacy ISO3 (ONLY for WGI / legacy joins)
    ISO3_host = case_when(
      ISO3_host_modern == "COD" ~ "ZAR",
      ISO3_host_modern == "ROU" ~ "ROM",
      TRUE ~ ISO3_host_modern
    ),
    ISO3_proponent = case_when(
      ISO3_proponent_modern == "COD" ~ "ZAR",
      ISO3_proponent_modern == "ROU" ~ "ROM",
      TRUE ~ ISO3_proponent_modern
    )
  )

##### Step 3 — Assign DEFAULT continents (do this ONCE) ####
# This preserves Africa, Europe, Asia, Oceania, etc.

Data_complete_iso <- Data_complete_iso %>%
  mutate(
    host_continent_iso = countrycode(
      ISO3_host_modern, "iso3c", "continent"
    ),
    proponent_continent_iso = countrycode(
      ISO3_proponent_modern, "iso3c", "continent"
    )
  )

##### Step 4 — Override ONLY the Americas ####
# This is the key step — nothing else is touched

Data_complete_iso <- Data_complete_iso %>%
  mutate(
    host_continent_iso = case_when(
      ISO3_host_modern %in% north_america_iso3   ~ "North America",
      ISO3_host_modern %in% central_america_iso3 ~ "Central America & Caribbean",
      ISO3_host_modern %in% south_america_iso3   ~ "South America",
      TRUE ~ host_continent_iso
    ),

    proponent_continent_iso = case_when(
      ISO3_proponent_modern %in% north_america_iso3   ~ "North America",
      ISO3_proponent_modern %in% central_america_iso3 ~ "Central America & Caribbean",
      ISO3_proponent_modern %in% south_america_iso3   ~ "South America",
      TRUE ~ proponent_continent_iso
    )
  )



# check
Data_complete_iso %>%
  count(host_continent_iso, sort = TRUE)




## check for continents

# No missing continents?
Data_complete_iso %>%
  summarise(
    missing_host_cont = sum(is.na(host_continent_iso)),
    missing_prop_cont = sum(is.na(proponent_continent_iso))
  )

# Which countries were legacy?
Data_complete_iso %>%
  filter(ISO3_host != ISO3_host_modern) %>%
  distinct(host_country_clean, ISO3_host_modern, ISO3_host)



# check for missing ISO codes

Data_complete_iso %>%
  summarise(
    missing_host_iso_modern = sum(is.na(ISO3_host_modern)),
    missing_prop_iso_modern = sum(is.na(ISO3_proponent_modern)),
    missing_host_iso_legacy = sum(is.na(ISO3_host)),
    missing_prop_iso_legacy = sum(is.na(ISO3_proponent))
  )


# check any rows failed to assign ISO

Data_complete_iso %>%
  filter(is.na(ISO3_host_modern) | is.na(ISO3_proponent_modern)) %>%
  distinct(
    host_country_or_area,
    proponent_country_manual,
    host_country_clean,
    proponent_country_clean
  )

## check where legacy ISO is used

Data_complete_iso %>%
  filter(ISO3_host != ISO3_host_modern) %>%
  distinct(host_country_clean, ISO3_host_modern, ISO3_host)


# check continent assign completeness

Data_complete_iso %>%
  summarise(
    missing_host_continent = sum(is.na(host_continent_iso)),
    missing_prop_continent = sum(is.na(proponent_continent_iso))
  )

# check continent values

unique(Data_complete_iso$host_continent_iso)
unique(Data_complete_iso$proponent_continent_iso)


# check for ISO collision

Data_complete_iso %>%
  distinct(ISO3_host_modern, host_country_clean) %>%
  count(ISO3_host_modern) %>%
  filter(n > 1)


## summary table

Data_complete_iso %>%
  summarise(
    n_rows = n(),
    n_unique_hosts = n_distinct(ISO3_host_modern),
    n_unique_proponents = n_distinct(ISO3_proponent_modern)
  )

save(
  Data_complete_iso,
  file = "ccbp_market_data_forest_based_iso.RDATA"
)




```

```{r}
#### PART 1 - Making governance plots for project host countries ####

# indicators include Voice and Accountability Political Stability and Absence of Violence/Terrorism, Government Effectiveness, Regulatory Quality, Rule of Law, Control of Corruption (https://www.worldbank.org/en/publication/worldwide-governance-indicators)

# GDP per capita data sourced from World Bank (https://data.worldbank.org/indicator/NY.GDP.PCAP.CD)



# This approach makes a facet wrap of all government indicators. Next we can use this to decide on indicators we care about, and what other variables we can include with graduated symbols or color scale #

library(tidyverse)
library(countrycode)
library(readxl)
library(scales)
library(ggrepel)
library(readr)

#### Step 1 - Sorting for only international traded projects ####

# starting with Data_complete_iso because it has already been cleaned, continents and ISO assigned, names cleaned in the above cleaning steps

Data_diff <- Data_complete_iso %>%
  filter(
    !is.na(ISO3_host_modern),
    !is.na(ISO3_proponent_modern),
    ISO3_host_modern != ISO3_proponent_modern
  )

# now we have reduced the dataset to only projects that have different proponent/hosts

# “We restricted analyses to international carbon projects, defined as projects where the buyer country differed from the host country, based on ISO-3 country codes after harmonising country name variants.”


#### Step 2 - Aggregate carbon volume by host country (ISO level) ####

host_carbon_totals <- Data_diff %>%
  group_by(ISO3_host_modern) %>%
  summarise(
    total_credit_volume = sum(sum_of_credit_volume, na.rm = TRUE),
    host_continent_iso = first(host_continent_iso),
    .groups = "drop"
  ) %>%
  rename(
    iso3c = ISO3_host_modern
  )


# check this

any(duplicated(host_carbon_totals$iso3c))




# should be FALSE

#### Step 3 - clean WGI data (governance metrics) ####

wgi_path <- "/Users/zoe/Desktop/CCBP/CCBP_database/data/gov_indicators_wgi/wgidataset.xlsx"

wgi_clean <- read_excel(wgi_path) %>%
  filter(
    indicator %in% c("cc", "ge", "pv", "rl", "rq", "va"),
    year == max(year, na.rm = TRUE)
  ) %>%
  mutate(
    estimate = parse_number(estimate)
  ) %>%
  select(
    iso3c = code,
    indicator,
    estimate
  )


#### Step 4 - pivot wider #####

wgi_wide <- wgi_clean %>%
  pivot_wider(
    names_from = indicator,
    values_from = estimate
  )



#### Step 5 - Join carbon data + governance metrics ####

host_carbon_wgi <- host_carbon_totals %>%
  left_join(wgi_wide, by = "iso3c")

# check this 

any(duplicated(wgi_wide$iso3c))
any(duplicated(host_carbon_totals$iso3c))

# should be FALSE


#### Step 6 - pivot longer for faceting ####

host_carbon_wgi_long <- host_carbon_wgi %>%
  pivot_longer(
    cols = c(cc, ge, pv, rl, rq, va),
    names_to = "indicator",
    values_to = "estimate"
  )


#### Step 7 - make indicator labels for gov. metrics ####

indicator_labels <- c(
  cc = "Control of Corruption",
  ge = "Government Effectiveness",
  pv = "Political Stability",
  rl = "Rule of Law",
  rq = "Regulatory Quality",
  va = "Voice & Accountability"
)

host_carbon_wgi_long <- host_carbon_wgi_long %>%
  mutate(
    indicator_label = indicator_labels[indicator]
  )


#### Step 8 - make country labels for each facet ####

label_countries_facets <- host_carbon_wgi_long %>%
  filter(
    total_credit_volume > 0,
    !is.na(estimate)
  ) %>%
  group_by(indicator, indicator_label) %>%
  summarise(
    data = list(
      bind_rows(
        slice_max(pick(everything()), total_credit_volume, n = 3),
        slice_min(pick(everything()), total_credit_volume, n = 3),
        slice_max(pick(everything()), estimate, n = 3),
        slice_min(pick(everything()), estimate, n = 3)
      ) %>%
        distinct(iso3c, .keep_all = TRUE)
    ),
    .groups = "drop"
  ) %>%
  unnest(data)

#### Step 9 - join GDP to carbon/WGI data for making graduated symbols ####

gdp <- read_csv(
  "/Users/zoe/Desktop/CCBP/CCBP_database/data/GDP_world_data/gdp_world_data_country.csv",
  show_col_types = FALSE
)

gdp_2024 <- gdp %>%
  transmute(
    iso3c = country_code,
    gdp_pc_2024 = as.numeric(`2024`)
  )

host_carbon_wgi_long <- host_carbon_wgi_long %>%
  left_join(gdp_2024, by = "iso3c")


#### Step 10 - Plot with GDP/continents/country labels ####

continent_colors <- c(
  "North America" = "#90C2E7",
  "Central America & Caribbean" = "#7088B4",
  "South America" = "#6D7F74",
  "Africa"   = "#4F4D80",
  "Asia" = "#A37466",
  "Europe" = "#98A886",
  "Oceania" = "#415662"
)

ggplot(
  host_carbon_wgi_long %>%
    filter(
      total_credit_volume > 0,
      !is.na(estimate),
      !is.na(gdp_pc_2024),
      !is.na(host_continent_iso)
    ),
  aes(
    x = total_credit_volume,
    y = estimate,
    size = gdp_pc_2024,
    colour = host_continent_iso
  )
) +
  geom_point(alpha = 0.7) +
  geom_text_repel(
    data = label_countries_facets,
    inherit.aes = FALSE,
    aes(
      x = total_credit_volume,
      y = estimate,
      label = iso3c
    ),
    size = 2.8,
    colour = "black",
    max.overlaps = Inf
  ) +
  scale_x_log10(
    labels = label_number(
      scale_cut = cut_si(""),
      accuracy = 1
    )
  ) +
  scale_size_continuous(
    name = "GDP per capita (USD, 2024)",
    range = c(1.5, 6),
    labels = label_dollar(accuracy = 1)
  ) +
  scale_colour_manual(
    name = "Continent",
    values = continent_colors
  ) +
  scale_y_continuous(n.breaks = 5) +
  facet_wrap(~ indicator_label, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Total forest-based carbon credit volume (log scale)",
    y = "Governance score (WGI)",
    title = "Forest-based carbon project volume vs governance, host countries",
    subtitle = "International projects only; point size = GDP per capita (2024)"
  ) +
  
  guides(
    colour = guide_legend(
      override.aes = list(size = 5)
    )
  )



#### Step 11 — Unlogged version (linear x-axis) ####

gov_host_plot_faceted_gdp_linear <- ggplot(
  host_carbon_wgi_long %>%
    filter(
      total_credit_volume > 0,
      !is.na(estimate),
      !is.na(gdp_pc_2024),
      !is.na(host_continent_iso)
    ),
  aes(
    x = total_credit_volume,
    y = estimate,
    size = gdp_pc_2024,
    colour = host_continent_iso
  )
) +
  geom_point(alpha = 0.7) +
  
  geom_text_repel(
    data = label_countries_facets,
    inherit.aes = FALSE,
    aes(
      x = total_credit_volume,
      y = estimate,
      label = iso3c
    ),
    size = 2.8,
    colour = "black",
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey60",
    max.overlaps = Inf
  ) +
  
  scale_x_continuous(
    labels = scales::label_number(
      scale_cut = scales::cut_si(""),
      accuracy = 1
    )
  ) +
  
  scale_size_continuous(
    name = "GDP per capita (USD, 2024)",
    range = c(1.5, 6),
    labels = scales::label_dollar(accuracy = 1)
  ) +
  
  scale_colour_manual(
    name = "Continent",
    values = continent_colors
  ) +
  
  scale_y_continuous(n.breaks = 5) +
  
  facet_wrap(~ indicator_label, scales = "free_y") +
  
  guides(
    colour = guide_legend(
      override.aes = list(size = 5)
    )
  ) +
  
  labs(
    x = "Total forest-based carbon credit volume",
    y = "Governance score (WGI)",
    title = "Forest-based carbon project volume vs governance, host countries",
    subtitle = "International projects only; linear x-axis; point size = GDP per capita (2024)"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "right"
  )

print(gov_host_plot_faceted_gdp_linear)


########### ------------------ ###############
########### ------------------ ###############
########### ------------------ ###############
########### ------------------ ###############
########### ------------------ ###############
########### ------------------ ###############

#### PART 2 - Making governance plots for project proponent countries ####


library(tidyverse)
library(countrycode)
library(readxl)
library(scales)
library(ggrepel)
library(readr)

#### Step 1 - PROPONENTS Sorting for only international traded projects ####

# starting with Data_complete_iso because it has already been cleaned, continents and ISO assigned, names cleaned in the above cleaning steps

Data_diff <- Data_complete_iso %>%
  filter(
    !is.na(ISO3_host_modern),
    !is.na(ISO3_proponent_modern),
    ISO3_host_modern != ISO3_proponent_modern
  )

# now we have reduced the dataset to only projects that have different proponent/hosts


#### Step 2 - Aggregate carbon volume by proponent country (ISO level) ####

proponent_carbon_totals <- Data_diff %>%
  group_by(ISO3_proponent_modern) %>%
  summarise(
    total_credit_volume = sum(sum_of_credit_volume, na.rm = TRUE),
    proponent_continent_iso = first(proponent_continent_iso),
    .groups = "drop"
  ) %>%
  rename(
    iso3c = ISO3_proponent_modern
  )


# check this

any(duplicated(proponent_carbon_totals$iso3c))

# should be FALSE

#### Step 3 - clean WGI data (governance metrics) ####

wgi_path <- "/Users/zoe/Desktop/CCBP/CCBP_database/data/gov_indicators_wgi/wgidataset.xlsx"

wgi_clean <- read_excel(wgi_path) %>%
  filter(
    indicator %in% c("cc", "ge", "pv", "rl", "rq", "va"),
    year == max(year, na.rm = TRUE)
  ) %>%
  mutate(
    estimate = parse_number(estimate)
  ) %>%
  select(
    iso3c = code,
    indicator,
    estimate
  )


#### Step 4 - pivot wider #####

wgi_wide <- wgi_clean %>%
  pivot_wider(
    names_from = indicator,
    values_from = estimate
  )



#### Step 5 - Join carbon data + governance metrics ####

proponent_carbon_wgi <- proponent_carbon_totals %>%
  left_join(wgi_wide, by = "iso3c")

# check this 

any(duplicated(wgi_wide$iso3c))
any(duplicated(proponent_carbon_totals$iso3c))

# should be FALSE


#### Step 6 - pivot longer for faceting ####

proponent_carbon_wgi_long <- proponent_carbon_wgi %>%
  pivot_longer(
    cols = c(cc, ge, pv, rl, rq, va),
    names_to = "indicator",
    values_to = "estimate"
  )


#### Step 7 - make indicator labels for gov. metrics ####

indicator_labels <- c(
  cc = "Control of Corruption",
  ge = "Government Effectiveness",
  pv = "Political Stability",
  rl = "Rule of Law",
  rq = "Regulatory Quality",
  va = "Voice & Accountability"
)

proponent_carbon_wgi_long <- proponent_carbon_wgi_long %>%
  mutate(
    indicator_label = indicator_labels[indicator]
  )


#### Step 8 - make country labels for each facet ####

label_countries_facets_proponents <- proponent_carbon_wgi_long %>%
  filter(
    total_credit_volume > 0,
    !is.na(estimate)
  ) %>%
  group_by(indicator, indicator_label) %>%
  summarise(
    data = list(
      bind_rows(
        slice_max(pick(everything()), total_credit_volume, n = 3),
        slice_min(pick(everything()), total_credit_volume, n = 3),
        slice_max(pick(everything()), estimate, n = 3),
        slice_min(pick(everything()), estimate, n = 3)
      ) %>%
        distinct(iso3c, .keep_all = TRUE)
    ),
    .groups = "drop"
  ) %>%
  unnest(data)

#### Step 9 - join GDP to carbon/WGI data for making graduated symbols ####

gdp <- read_csv(
  "/Users/zoe/Desktop/CCBP/CCBP_database/data/GDP_world_data/gdp_world_data_country.csv",
  show_col_types = FALSE
)

gdp_2024 <- gdp %>%
  transmute(
    iso3c = country_code,
    gdp_pc_2024 = as.numeric(`2024`)
  )

proponent_carbon_wgi_long <- proponent_carbon_wgi_long %>%
  left_join(gdp_2024, by = "iso3c")


#### Step 10 - Plot with GDP/continents/country labels ####

continent_colors <- c(
  "North America" = "#90C2E7",
  "Central America & Caribbean" = "#7088B4",
  "South America" = "#6D7F74",
  "Africa"   = "#4F4D80",
  "Asia" = "#A37466",
  "Europe" = "#98A886",
  "Oceania" = "#415662"
)

gov_proponent_plot_faceted_gdp_logged <- ggplot(
  proponent_carbon_wgi_long %>%
    filter(
      total_credit_volume > 0,
      !is.na(estimate),
      !is.na(gdp_pc_2024),
      !is.na(proponent_continent_iso)
    ),
  aes(
    x = total_credit_volume,
    y = estimate,
    size = gdp_pc_2024,
    colour = proponent_continent_iso
  )
) +
  geom_point(alpha = 0.7) +
  geom_text_repel(
    data = label_countries_facets_proponents,
    inherit.aes = FALSE,
    aes(
      x = total_credit_volume,
      y = estimate,
      label = iso3c
    ),
    size = 2.8,
    colour = "black",
    max.overlaps = Inf
  ) +
  scale_x_log10(
    labels = label_number(
      scale_cut = cut_si(""),
      accuracy = 1
    )
  ) +
  scale_size_continuous(
    name = "GDP per capita (USD, 2024)",
    range = c(1.5, 6),
    labels = label_dollar(accuracy = 1)
  ) +
  scale_colour_manual(
    name = "Continent",
    values = continent_colors
  ) +
  scale_y_continuous(n.breaks = 5) +
  facet_wrap(~ indicator_label, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Total forest-based carbon credit volume (log scale)",
    y = "Governance score (WGI)",
    title = "Forest-based carbon project volume vs governance, proponent countries",
    subtitle = "International projects only; point size = GDP per capita (2024)"
  ) +
  
  guides(
    colour = guide_legend(
      override.aes = list(size = 5)
    )
  )

gov_proponent_plot_faceted_gdp_logged

#### Step 11 — Unlogged version (linear x-axis) ####

gov_proponent_plot_faceted_gdp_linear <- ggplot(
  proponent_carbon_wgi_long %>%
    filter(
      total_credit_volume > 0,
      !is.na(estimate),
      !is.na(gdp_pc_2024),
      !is.na(proponent_continent_iso)
    ),
  aes(
    x = total_credit_volume,
    y = estimate,
    size = gdp_pc_2024,
    colour = proponent_continent_iso
  )
) +
  geom_point(alpha = 0.7) +
  
  geom_text_repel(
    data = label_countries_facets_proponents,
    inherit.aes = FALSE,
    aes(
      x = total_credit_volume,
      y = estimate,
      label = iso3c
    ),
    size = 2.8,
    colour = "black",
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "grey60",
    max.overlaps = Inf
  ) +
  
  scale_x_continuous(
    labels = scales::label_number(
      scale_cut = scales::cut_si(""),
      accuracy = 1
    )
  ) +
  
  scale_size_continuous(
    name = "GDP per capita (USD, 2024)",
    range = c(1.5, 6),
    labels = scales::label_dollar(accuracy = 1)
  ) +
  
  scale_colour_manual(
    name = "Continent",
    values = continent_colors
  ) +
  
  scale_y_continuous(n.breaks = 5) +
  
  facet_wrap(~ indicator_label, scales = "free_y") +
  
  guides(
    colour = guide_legend(
      override.aes = list(size = 5)
    )
  ) +
  
  labs(
    x = "Total forest-based carbon credit volume",
    y = "Governance score (WGI)",
    title = "Forest-based carbon project volume vs governance, proponent countries",
    subtitle = "International projects only; linear x-axis; point size = GDP per capita (2024)"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "right"
  )

print(gov_proponent_plot_faceted_gdp_linear)



##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####
##### ---- #####




####### Plotting governance metrics comparing host and proponents ########

##### HOST AND PROPONENT PLOT - with RQ gov metric only ####

#### Step 1 - prepare RQ for hosts and props ####

host_rq <- host_carbon_wgi_long %>%
filter(
  indicator == "rq",
  total_credit_volume > 0,
  !is.na(estimate)
) %>%
  mutate(role = "Host")

proponent_rq <- proponent_carbon_wgi_long %>%
  filter(
    indicator == "rq",
    total_credit_volume > 0,
    !is.na(estimate)
  ) %>%
  mutate(role = "Proponent")


#### Step 2 - combine into dataframe #####

rq_combined <- bind_rows(host_rq, proponent_rq)

#check
table(rq_combined$role)

#### Step 3 - plot ####

rq_overlay_plot <- ggplot(
  rq_combined,
  aes(
    x = total_credit_volume,
    y = estimate,
    colour = role
  )
) +
  geom_point(
    alpha = 0.6,
    size = 2
  ) +
  
  scale_colour_manual(
    values = c(
      "Host" = "#8ECae6",      # light blue
      "Proponent" = "#CDB4DB"  # light purple
    ),
    name = "Country role"
  ) +
  
  scale_x_log10(
    labels = label_number(
      scale_cut = cut_si(""),
      accuracy = 1
    )
  ) +
  
  scale_y_continuous(n.breaks = 5) +
  
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    colour = "grey60"
  ) +
  
  labs(
    x = "Total forest-based carbon credit volume (log scale)",
    y = "Regulatory Quality (WGI)",
    title = "Regulatory quality vs forest-based carbon credit volume",
    subtitle = "Host countries (blue) vs proponent countries (purple)"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )

print(rq_overlay_plot)

#### UNLOGGED #####

rq_overlay_plot_linear <- ggplot(
  rq_combined,
  aes(
    x = total_credit_volume,
    y = estimate,
    colour = role
  )
) +
  geom_point(
    alpha = 0.6,
    size = 2
  ) +
  
  scale_colour_manual(
    values = c(
      "Host" = "#17bebb",      # light blue
      "Proponent" = "#cd5334"  # light purple
    ),
    name = "Country role"
  ) +
  
  scale_x_continuous(
    labels = scales::label_number(
      scale_cut = scales::cut_si(""),
      accuracy = 1
    )
  ) +
  
  scale_y_continuous(n.breaks = 5) +
  
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    colour = "grey60"
  ) +
  
  labs(
    x = "Total forest-based carbon credit volume",
    y = "Regulatory Quality (WGI)",
    title = "Regulatory quality vs forest-based carbon credit volume",
    subtitle = "Host countries (blue) vs proponent countries (purple); linear scale"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )

print(rq_overlay_plot_linear)

















```

```



```{r}
#### preparing data for alluvial plot #####

Data_diff <- Data_complete %>%
  filter(proponent_country_manual != host_country_or_area)

flow_matrix <- Data_diff %>%
  mutate(sum_of_credit_volume = as.numeric(sum_of_credit_volume)) %>%  # Convert to numeric
  group_by(buyer_continent, project_continent) %>%
  summarise(total_volume = sum(sum_of_credit_volume, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = project_continent,
    values_from = total_volume,
    values_fill = 0
  ) %>%
  column_to_rownames("buyer_continent")

# View the result
print(flow_matrix)

#Put it into alluvial format
# Convert the matrix to a data frame
flow_df <- as.data.frame(flow_matrix)

# Add the source column from row names
flow_df <- flow_df %>%
  mutate(source = rownames(flow_matrix)) %>%
  relocate(source)

# View the result
print(flow_df)

df_long <- flow_df %>%
  pivot_longer(cols = -source, names_to = "target", values_to = "value") %>%
  filter(value > 0) %>%  # Optional: remove zero flows
  mutate(
    source = factor(source, levels = unique(source)),
    target = factor(target, levels = unique(target))
  )

#Create alluvial plots for flows between regions
#Plot with labels
p <- ggplot(df_long,
       aes(axis1 = source, axis2 = target, y = value)) +
  geom_alluvium(aes(fill = source), width = 1/12) +
  geom_stratum(width = 1/12, fill = "grey", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2) +
  scale_x_discrete(limits = c("Origin", "Retirement"), expand = c(.05, .05))  +
  scale_fill_manual(values = c(
    "Africa" = "#4f2fee",
    "Asia" = "#42999b",
    "Europe" = "#70ffc4",
    "North America" = "#fac723",
    "South America" = "#ff5c9a",
    "Oceania" = "#b25b68"
  )) +
  theme_minimal() +
  ggtitle("Forest-based carbon flows between regions")

#Plot without labels
p2 <- ggplot(df_long,
       aes(axis1 = source, axis2 = target, y = value)) +
  geom_alluvium(aes(fill = source), width = 1/12) +
  geom_stratum(width = 1/12, fill = "grey", color = "black") +
  #geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2) +
  scale_x_discrete(limits = c("Origin", "Retirement"), expand = c(.05, .05))  +
  scale_fill_manual(values = c(
    "Africa" = "#4f2fee",
    "Asia" = "#42999b",
    "Europe" = "#70ffc4",
    "North America" = "#fac723",
    "South America" = "#ff5c9a",
    "Oceania" = "#b25b68"
  )) +
  theme_minimal() +
  ggtitle("Forest-based carbon flows between regions")

print(p2)
p2

# Export to both plots
ggsave("/Users/zoe/Desktop/CCBP/CCBP_database/figures/alluvial_plot_text_zl_testing.png", plot = p, width = 10, height = 6, dpi = 300)
ggsave("/Users/zoe/Desktop/CCBP/CCBP_database/figures/alluvial_plot_zl_testing.png", plot = p2, width = 10, height = 6, dpi = 300)

#Create the plot for flows between countries
# Make sure Credit.Volume is numeric
Data_complete <- Data_diff %>%
  mutate(Credit.Volume = as.numeric(gsub(",", "", Credit.Volume)))

# Summarise and reshape to long format for alluvial plot
df_long_country <- Data_complete %>%
  group_by(Headquarter.Country, Project.Country) %>%
  summarise(value = sum(Credit.Volume, na.rm = TRUE), .groups = "drop") %>%
  rename(source = Headquarter.Country, target = Project.Country)

# Create the alluvial plot with labels
p3 <- ggplot(df_long_country,
            aes(axis1 = source, axis2 = target, y = value)) +
  geom_alluvium(aes(fill = source), width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey90", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, hjust = 0) +
  scale_x_discrete(limits = c("Origin", "Retirement"), expand = c(.05, .05)) +
  theme_minimal() +
  ggtitle("Forest-based carbon flows between countries")

# Print the plot
print(p3)
ggsave("C:/Users/uqbwil14/Dropbox/Carbon crossing borders project/Rcode_bw/CCBP/figures/countries_text.png", plot = p3, width = 10, height = 6, dpi = 300)

# Create the alluvial plot without labels
p4 <- ggplot(df_long_country,
            aes(axis1 = source, axis2 = target, y = value)) +
  geom_alluvium(aes(fill = source), width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/12, fill = "grey90", color = "black") +
  #geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 2, hjust = 0) +
  scale_x_discrete(limits = c("Origin", "Retirement"), expand = c(.05, .05)) +
  theme_minimal() +
  ggtitle("Forest-based carbon flows between countries")


# Print the plot
print(p4)
ggsave("C:/Users/uqbwil14/Dropbox/Carbon crossing borders project/Rcode_bw/CCBP/figures/countries.png", plot = p4, width = 10, height = 6, dpi = 300)

```

Summaries for reporting purposes
```{r}
# Make sure Credit.Volume is numeric
Data_complete <- Data_complete %>%
  mutate(Credit.Volume = as.numeric(Credit.Volume))

# Group and summarise
top_countries <- Data_complete %>%
  group_by(Headquarter.Country) %>%
  summarise(Total_Credit = sum(Credit.Volume, na.rm = TRUE)) %>%
  arrange(desc(Total_Credit))

# View result
print(top_countries)

# Summarise by Project.Country
top_projects <- Data_complete %>%
  group_by(Project.Country) %>%
  summarise(Total_Credit = sum(Credit.Volume, na.rm = TRUE)) %>%
  arrange(desc(Total_Credit)) %>%
  slice_head(n = 10)  # Top 10

# View result
print(top_projects)

## hello ##

print(top_projects)
```

```{r}
verra_projects <- subset(Data, Registry == "Verra")
unique(verra_projects$Project.Methodologies)

```
```{r}

